---
title: "Some Personal Projects"
author: "John Goldin"
date: "3/16/2021"
output: 
  html_document:
    theme: flatly
    toc: TRUE
    toc_depth: 3
    toc_float: 
      collapsed: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<META NAME="ROBOTS" CONTENT="NOINDEX, NOFOLLOW">
## Notes to Accompany My Talk on Personal Projects

### Having a Presence on the Web

Me: [johngoldin.com](https://www.johngoldin.com)

I am very R oriented so I'm using RMarkdown, blogdown, and
web hosting via [Netlify](https://www.netlify.com). Years ago I had a
[WordPress](https://wordpress.com) blog, but I don't
know anything about that world these days.

For a good introduction to publishing to the web using RStudio
and RMarkdown, see 
[How to Get Your Materials Online With R Markdown](https://www.youtube.com/watch?v=QcE4RBH2auQ)
by Alison Hill and Desirée De Leon.
Their tutorial shows several techniques:

- a single RMarkdown document as a web page

- a simple RMarkdown web site (at 8:00 in the video)

- using the RStudio package [distill](https://rstudio.github.io/distill/) to publish a simple web site (at 17:09 on the video)

- using [bookdown](https://www.bookdown.org) to publish a book (e.g., [R for Data Science](https://r4ds.had.co.nz)) (at 23:46 on the video)

- using [blogdown](https://bookdown.org/yihui/blogdown/) and 
[Hugo](https://gohugo.io) to create a blog site (at 34:03 on the video)

  For a full-featured blog using Hugo, see Alison Hill's post [Up and Running with blogdown](https://alison.rbind.io/post/2017-06-12-up-and-running-with-blogdown/) 

This web page was deployed via the first (and simplest) technique described in their video. 
I dragged
the project folder containing the RMarkdown html code and dropped it on the Netlify page
to deploy it. Initially it had an arbitrary name created by Netlify, 
but I was able to change it to
[https://goldin-projects-2021.netlify.app/](https://goldin-projects-2021.netlify.app/).

At the top of the RMarkdown file that created this page, I included the
following line:

  `<META NAME="ROBOTS" CONTENT="NOINDEX, NOFOLLOW">`

This tells search engines to skip over this page and should keep it out of search
results. That might be useful if you have created a lighthearted page that you
would rather not turn up when you apply for some position in the future 
(or run for president).

### Mining Publicly Available Data via Web API's

In the R world there are many packages that provide functions to facilitate
using particular Web API's. For example, [tidycensus](https://walker-data.com/tidycensus/) 
and [tigris](https://github.com/walkerke/tigris) make it 
dramatically easier to access US Census data, especially the American Community
Survey. There's also [censusxy](https://cran.r-project.org/web/packages/censusxy/vignettes/censusxy.html)
for geocoding from the Census Bureau. I did a post which I think is a pretty
good example of a [novice using tidycensus to explore census data](https://www.johngoldin.com/2019/04/new-haven-census-and-r/).

A number of government entities use a commercial package called [Socrata](https://www.tylertech.com/products/socrata/open-data-citizen-engagement)
to make data available to the public. I use the [RSocrata](https://github.com/Chicago/RSocrata) package to retrieve
data both from [data.ct.gov](https://data.ct.gov) and [data.cdc.gov](https://data.cdc.gov). 

On another occasion I had fun playing
with Twitter data fetched via the [twitteR](https://cran.r-project.org/web/packages/twitteR/twitteR.pdf) or [rtweet](https://docs.ropensci.org/rtweet/) packages. Both access the 
Twitter API. 

A widely used source of economic data is the [FRED database](https://fred.stlouisfed.org/), which in R
can be accessed via the [fredr](http://sboysel.github.io/fredr/articles/fredr.html)
package. I suspect that to get 
[Current Population Survey Data](https://fred.stlouisfed.org/categories/12) 
one could do better
with FRED than with the source agencies (Census Bureau and BLS). The same
is probably true for many other measures. As an example of the wider world of API's
beyond R, see their list of 
[FRED toolkits](https://fred.stlouisfed.org/docs/api/fred/#Toolkits)
for many computer languages.

If an R package is not available, it may still be possible to use a particular
API. For example, I was able to follow an example presented by [Hadley Wickham]([examples](https://github.com/hadley/vis-eda/blob/master/vis-eda.pdf) to
use the TripIt API to retrieve details of my travels that were stored in TripIt.
I used flight records to identify when I was out of the Eastern Time Zone so that
I could [adjust timestamps](https://www.johngoldin.com/2020/02/apple-health-export-part-i/) 
of data that I retrieved from the Apple Health Export. 
One of my early efforts was to [retrieve photos I stored in Flickr](https://www.johngoldin.com/2016/07/flickr_api/) via the Flickr API.

When you are using an API directly (i.e., not via an R package) you may
quickly end up lost amidst technical details. 
One may hear terms like [REST or SOAP](https://raygun.com/blog/soap-vs-rest-vs-json/) that seem mysterious.
Or you may have to cope with data formats such as JSON or XML. I've had success with the [jsonlite](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html) package
to fetch JSON data. XML data is more of a challenge for me, and there are several
XML related R packages.

```{r libd, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(lubridate)
library(kableExtra)
library(janitor)
library(knitr)
library(scales)
path_saved_export <- "~/Dropbox/Programming/R_Stuff/john_vitals/Apple-Health-Data/"
path_to_healthexport1 <- "~/Documents/R_local_repos/applehealth1/R/"
```
```{r load_health_df, message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE}
load(paste0(path_saved_export,"save_processed_export.RData"))
```
### Apple Health Export

#### Key R Code That Imports the Apple Health Export

```{r eval = FALSE}
library(tidyverse)
library(XML)
unzip("~/Downloads/export.zip", exdir = "~/Downloads", overwrite = TRUE)
# Note that when unzipped, the resulting folder can be HUGE!'
# In my case, over a gigabyte. (XML is a verbose format.) I always
# delete the folder right after I import it.
health_xml <- xmlParse("~/Downloads/apple_health_export/export.xml")
# takes about 70 seconds on my iMac
health_df <- XML:::xmlAttrsToDataFrame(health_xml["//Record"], 
                                       stringsAsFactors = FALSE) %>%
  as_tibble() %>% mutate(value = as.numeric(value))
activity_df <- XML:::xmlAttrsToDataFrame(health_xml["//ActivitySummary"], 
                                         stringsAsFactors = FALSE) %>% 
  as_tibble()
workout_df <-  XML:::xmlAttrsToDataFrame(health_xml["//Workout"], 
                                         stringsAsFactors = FALSE) %>% 
  as_tibble
clinical_df <- XML:::xmlAttrsToDataFrame(health_xml["//ClinicalRecord"]) %>% 
  as_tibble()
```

### Frequency of Data by Type

Data from Apple Watch since October 2017, iPhone since December 2014.    

```{r type_table, echo = FALSE}

health_df %>% 
  tabyl(type) %>% 
  arrange(desc(n)) %>% 
  janitor::adorn_totals("row") %>% # do column total after arrange
  mutate(percent = round(percent, 3),
    percent = scales::percent(percent)) %>% 
  kable(format.args = list(decimal.mark = " ", big.mark = ","),
        align = c("lrr"),
        # table.attr='class="myTable"', 
        caption = "Frequency Data by Type", format = "html") %>% 
  kable_styling()
        # caption = "Frequency Data by Category")
```
