---
title: "Personal Projects"
author: 
  - "John Goldin"
date: '`r format(Sys.Date(),"%A, %B %d")`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%" 
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

<br><br>
### Motto for this talk:
<br>

# What can you do when you're retired...

--

# ...anything you want

---
### Topics

### Having a personal presence on the web

### An abundance of data via API's

### My current deep dive into the Apple Health Export

---
## Having a Presence on the Web
### Why?

####"Publication precipitates reality." -- Lila Freedman, former editor of the Yale Blue Book
#####My blog: [johngoldin.com](https://www.johngoldin.com)
--

####Feeling part of a larger world
+ Google [American walker in Britain](https://www.google.com/search?q=American+walker+in+britain)
+ Google ["Apple Health Export"](https://www.google.com/search?q=%22apple+health+export%22)
--

####Professional visibility (not so relevant for a retiree, and maybe a risk for others)
--

###Caveats:

####Web is very public: most of your Yale work is, at best, semi-private

---
## Publishing to the Web

I've created some notes to accompany this talk. They are available at    

https://goldin-projects-2021.netlify.app/

The notes point you to a video by Alison Hill & Desirée De Leon that describe a
number of ways to use RMarkdown to publish to the web. The first and simplest
technique was used to create the notes page. 

1. Knit an RMarkdown document 
2. Sign into Netlify.com 
1. Drag the folder containing the project onto the deploy
page in Netlify

When first deployed, Netlify creates an
arbitray name for the URL. But I have the option to change it,
provided I can supply a name that hasn't already been used.

There are lots of other ways to have a presence on the Web,
such as LinkedIn, Twitter, Instagram, Flickr, and so on.

---
## Where To Find Some Data

### Getting data via web API's

+ Connecticut and CDC [Covid-19 data](https://data.ct.gov/browse?tags=covid-19) (via RSocrata)

```{r socrata_example, eval = FALSE, echo = TRUE}
dph_counties <- read.socrata("https://data.ct.gov/resource/bfnu-rgqt.json",
                            app_token = Sys.getenv("CTDATA_APP1_TOKEN")) %>%
  as_tibble() %>%
  rename(cases = totalcases, deaths = totaldeaths) %>%
  mutate(date = as_date(dateupdated),
         cases = as.numeric(cases), deaths = as.numeric(deaths),
         tests = NA_real_, tests_positive = NA_real_,
         hospital = as.numeric(hospitalization),
         confirmedcases = as.numeric(confirmedcases),
         confirmeddeaths = as.numeric(confirmeddeaths),
         probablecases = as.numeric(probablecases),
         probabledeaths = as.numeric(probabledeaths)) %>%
  select(-dateupdated, -hospitalization)
```

---
## An API is much easier to use if there is an R package that manages it

Some examples:

US Census: [tidycensus](https://walker-data.com/tidycensus/) 
and [tigris](https://github.com/walkerke/tigris)

Twitter: [twitteR](https://cran.r-project.org/web/packages/twitteR/twitteR.pdf) or [rtweet](https://docs.ropensci.org/rtweet/)

[FRED](https://fred.stlouisfed.org/) economic stats: [fredr](http://sboysel.github.io/fredr/articles/fredr.html)


More links in the notes at [goldin-projects-2021.netlify.app](https://goldin-projects-2021.netlify.app/)

```{r libd, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(lubridate)
library(kableExtra)
library(janitor)
library(scales)
path_saved_export <- "~/Dropbox/Programming/R_Stuff/john_vitals/Apple-Health-Data/"
path_to_healthexport1 <- "~/Documents/R_local_repos/applehealth1/R/"
```
```{r load_health_df, message = FALSE, warning = FALSE, echo = FALSE, cache = TRUE}
load(paste0(path_saved_export,"save_processed_export.RData"))
```

---
#### I don't need no stinkin' guvment databases to satisfy my data anlysis habit.

#### I've got almost three million rows of data in my pocket!

--

I did a [blog post](https://johngoldin.com/2020/02/apple-health-export-part-i/) 
last winter describes how to export Apple Health Export and load it into R.

---
### What's in the Apple Health Export dataset?

.pull-left[
```{r basic_table, echo = FALSE, message = TRUE}
for_table <- health_df %>%
  tabyl(category) %>%
  arrange(desc(n)) %>%
  janitor::adorn_totals("row") %>% # do column total after arrange
  mutate(percent = scales::percent(round(percent, 2), 1))
  # janitor::adorn_totals("col") %>%  arrange(desc(Total)) %>%
```
```{r basic_table-left, echo = FALSE, message = TRUE}
  kable(for_table[1:10,], format.args = list(decimal.mark = " ", big.mark = ","),
        align = c("lrr"),
        # table.attr='class="myTable"',
        caption = "Frequency Data by Category", format = "html") %>%
  kable_styling()
        # caption = "Frequency Data by Category")
```

]
.pull-right[ 
```{r basic_table-right, echo = FALSE, message = TRUE}
  # had to explicitly convert tabyl to tibble if not starting from row 1
  kable(for_table[11:nrow(for_table),] %>% as_tibble(), format.args = list(decimal.mark = " ", big.mark = ",", row.names = FALSE),
        align = c("lrr"),
        # table.attr='class="myTable"',
        caption = "Frequency Data by Category", format = "html") %>%
  kable_styling()
        # caption = "Frequency by Category of Data")
```
]

<br>
More detail about types of data shown in the notes at [goldin-projects-2021.netlify.app](https://goldin-projects-2021.netlify.app/)

``` 


```{r hr_boundaries, echo = FALSE, message = FALSE, cache = TRUE}
hr <- health_df %>% 
  filter(type == "RestingHeartRate")

hr_versions <- unique(hr %>% filter(str_detect(sourceName, "Watch"), !is.na(sourceVersion)) %>% 
                     select(sourceVersion)) %>% 
                     separate(sourceVersion, into = c("major", "minor", "subminor"), remove = FALSE) %>% 
  mutate(subminor = ifelse(is.na(subminor), "0", subminor)) %>% 
  arrange(sourceVersion)

hr_boundaries <- hr %>% 
  left_join(hr_versions, by = "sourceVersion") %>% 
  arrange(sourceVersion, local_start, creationDate) %>% 
  group_by(sourceVersion, major, minor, subminor) %>% 
  summarise(first_date = first(local_start), last_date = last(local_start))

hr_boundaries <- hr_boundaries %>% 
  ungroup() %>% 
  arrange(first_date) %>% 
  mutate(test = lag(major),
         test2 = lead(major),
    level = case_when(
    major != lag(major) ~ 3,
    minor != lag(minor) ~ 2, 
    subminor != lag(subminor) ~ 1,
    TRUE ~ NA_real_
  ))
p_noversion1 <- ggplot(data = hr, aes(x = local_date, y = value)) +
  ylab("resting heart rate") + xlab(NULL) +
  ggtitle("Resting Heart Rate by Day") +
  geom_point(size = 0.5) 
p_noversion2 <- p_noversion1 +
  geom_smooth()
p_version1 <- p_noversion1 +
  geom_vline(data = hr_boundaries %>% filter(level == 3), 
             aes(xintercept = as_date(first_date)), size = 1, colour = "seagreen3") +
  geom_vline(data = hr_boundaries %>% filter(level == 2), 
             aes(xintercept = as_date(first_date)), size = 0.3, colour = "seagreen2") 
big_change <- hr_boundaries$first_date[hr_boundaries$sourceVersion == "7.0"] %>% as_date()
p_version2 <- p_version1 +
  geom_smooth(data = hr %>% filter(local_date < big_change)) +
  geom_smooth(data = hr %>% filter(local_date >= big_change)) 

```
```{r vo2max, echo = FALSE, message = FALSE, cache = TRUE}
vo2max <- health_df %>% 
  filter(type == "VO2Max")
vo_versions <- unique(vo2max %>% filter(type == "VO2Max", !is.na(sourceVersion)) %>% 
                     select(sourceVersion)) %>% 
                     separate(sourceVersion, into = c("major", "minor", "subminor"), remove = FALSE) %>% 
  mutate(subminor = ifelse(is.na(subminor), "0", subminor), major = as.numeric(major),
         minor = as.numeric(minor), subminor = as.numeric(subminor)) %>% 
  arrange(major, minor, subminor)
vo2max2 <- vo2max %>% 
  left_join(vo_versions, by = "sourceVersion") %>% 
  arrange(sourceVersion, local_start, creationDate)
vo_boundaries <- vo2max2 %>% 
  group_by(major, minor, subminor) %>% 
  summarise(first_date = first(local_start), last_date = last(local_start))

vo_boundaries <- vo_boundaries %>% 
  ungroup() %>% 
  arrange(first_date) %>% 
  mutate(test = lag(major),
         test2 = lead(major),
    level = case_when(
    (major != lag(major)) | ((major > 0) & is.na(lag(major))) ~ 3,
    is.na(lag(major)) ~ 0,
    minor != lag(minor) ~ 2,
    subminor != lag(subminor) ~ 1,
    TRUE ~ NA_real_
  ))
p_vo2max1 <- ggplot(data = vo2max, aes(x = local_date, y = value)) +
  ylab("VO2 Max") + xlab(NULL) +
  ggtitle("Estimated VO2 Max") +
  geom_point(size = 0.5) 
p_vo2max2 <- p_vo2max1 # no smooth
p_vo2max3 <- p_vo2max2 +
  geom_vline(data = vo_boundaries %>% filter(level == 3), 
             aes(xintercept = as_date(first_date)), size = 1, colour = "goldenrod1") +
  geom_vline(data = vo_boundaries %>% filter(level == 2), 
             aes(xintercept = as_date(first_date)), size = 0.3, colour = "goldenrod3") +
  geom_vline(data = hr_boundaries %>% filter(level == 3), 
             aes(xintercept = as_date(first_date)), size = 1, colour = "seagreen3") +
  geom_vline(data = hr_boundaries %>% filter(level == 2), 
             aes(xintercept = as_date(first_date)), size = 0.3, colour = "seagreen2") 
```

---
```{r, echo = FALSE, message = FALSE}
print(p_noversion1)
```

---
```{r, echo = FALSE, message = FALSE}
print(p_noversion2)
```

---
```{r, echo = FALSE, message = FALSE}
print(p_version1)
```
---
```{r, echo = FALSE, message = FALSE}
print(p_version2)
```


---
```{r , echo = FALSE, message = FALSE}
print(p_vo2max1)
```

---
```{r , echo = FALSE, message = FALSE}
print(p_vo2max3)
```


