---
title: "Personal Projects"
author: "John Goldin"
date: "3/16/2021"
output: 
  ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Motto for this presentation {.build}

What can you do when you're retired...

...anything you want

## Topics

Having a personal presence on the web

An abundance of data via API's

My current deep dive into the Apple Health Export

## Having a Presence on the Web {.build}

Why?

+ "Publication precipitates reality." -- Lila Freedman, former editor of the Yale Blue Book

+ Feeling part of a larger world

  + Google [American walker in Britain](https://www.google.com/search?q=American+walker+in+britain)
  
  + Google ["Apple Health Export"](https://www.google.com/search?q=%22apple+health+export%22)

+ Professional visibility (not so relevant for a retiree, and maybe a double-edged sword)
  
Me: [johngoldin.com](https://www.johngoldin.com)

Caveats:

+ Web is very public: most of your Yale work is, at best, semi-private

## Publishing to the Web

I've created some notes to accompany this talk. They are available at

https://goldin-projects-2021.netlify.app/

The notes point you to a video by Alison Hill & Desirée De Leon that describe a
number of ways to use RMarkdown to publish to the web. The first and simplest
technique was used to create the notes page.

## Where To Find Some Data

### Getting data via web API's

+ Connecticut and CDC [Covid-19 data](https://data.ct.gov/browse?tags=covid-19) (via RSocrata)

```{r socrata_example, eval = FALSE, echo = TRUE}
dph_counties <- read.socrata("https://data.ct.gov/resource/bfnu-rgqt.json",
                            app_token = Sys.getenv("CTDATA_APP1_TOKEN")) %>%
  as_tibble() %>%
  rename(cases = totalcases, deaths = totaldeaths) %>%
  mutate(date = as_date(dateupdated),
         cases = as.numeric(cases), deaths = as.numeric(deaths),
         tests = NA_real_, tests_positive = NA_real_,
         hospital = as.numeric(hospitalization),
         confirmedcases = as.numeric(confirmedcases),
         confirmeddeaths = as.numeric(confirmeddeaths),
         probablecases = as.numeric(probablecases),
         probabledeaths = as.numeric(probabledeaths)) %>%
  select(-dateupdated, -hospitalization)
```

## An API is much easier to use if there is an R package that manages it

Some examples:

US Census: [tidycensus](https://walker-data.com/tidycensus/) 
and [tigris](https://github.com/walkerke/tigris)

Twitter: [twitteR](https://cran.r-project.org/web/packages/twitteR/twitteR.pdf) or [rtweet](https://docs.ropensci.org/rtweet/)

[FRED](https://fred.stlouisfed.org/) economic stats: [fredr](http://sboysel.github.io/fredr/articles/fredr.html)
